### 2010-05-24 paper/title.tex
switched to image of both planes, respaced so the image could be bigger

\begin{titlepage}
\begin{center}

\includegraphics[width=0.6\textwidth]{../images/bothplanes_isometric.jpg}\\[7mm]
\HRule \\[7mm]
{ \huge \bfseries Rutgers University Autonomous Aircraft Team } \\[0.4cm]
\HRule \\[0.4cm]
{ \large \bfseries Technical Report }
\\[0.4cm]
{ \large \bfseries 2010 AUVSI UAS Competition }
\\[0.3cm]
\HRule \\[7mm]

  {\large Amanda Gaetano, Anthony Garrison, Pat Hickey,}
\\{\large Stephen Indyk, Bradley Lord, Cogan Noll,}
\\{\large John Palmer, Adrien Perkins, Gregory Quinn,}
\\{\large Michael Varga, Taras Wallace, and David Wescott}
\\[1cm]
\emph{Thank you to our sponsors:}
\\ Rutgers University Engineering Governance Council
\\ Rutgers Alumni Association
\\ WINLAB
\\ Invensense, Inc.
\\ ST Micro, Inc.
\\ BP Hobbies

\vfill
{\large \today}

\end{center}
\end{titlepage}